---
title: "Building Bioconductor workshops workflow"
author:
- name: Marcel Ramos
  affiliation: Roswell Park Cancer Institute, Buffalo, NY
date: "`r format(Sys.Date(), '%A, %B %d, %Y')`"
vignette: |
  %\VignetteIndexEntry{workshopbuilder}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
output:
  BiocStyle::html_document:
    toc_float: true
Package: workshopbuilder
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

# Installation

```{r}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("Bioconductor/workshopbuilder")
```

```{r,eval=TRUE,echo=FALSE}
.options <- workshopbuilder:::.options
```

# Set up

`workshopbuilder` ships with a number of package options to make management
of workshop directories and repositories easier. The three main options are:

1) Workshop book repository location on GitHub


```{r, eval=TRUE}
.options$set("BOOK_REPO", "Bioconductor/BiocWorkshops2019")
```

2) Local workshop book directory

```{r, eval=TRUE}
.options$set("LOCAL_REPO", normalizePath("~/BiocWorkshops2019"))
```

3) Local workshop repositories directory

```{r, eval=TRUE}
.options$set("REPOS_PATH", normalizePath("~/Bioconductor/Workshops"))
```

# Workflow Overview

The `workshopbuilder` package facilitates the installation and rendering
of the workshop book repository along with all the component repositories.

The individual workshop repositories are documented as GitHub issues in
the main workshop book repository.

We then take the workshops and the main book repository and clone them all.
We install and transfer vignettes to the main book repository and run a
render step to generate the book document.

## Read the issues from the repository issues comments

`workshopbuilder` takes issues with a `[Workshop]` tag and pulls workshops
from the URL provided in the body of the issue.

## Clone the workshop repositories (`pull` if already available)

The package uses `git2r` to clone all invididual workshop repositories
(or those otherwise indicated) to the location as specified by the `REPOS_PATH`
package option (see setup).


## Install the workshops

```{r,eval=TRUE,include=TRUE,results="hide",message=FALSE,warning=FALSE}
library(workshopbuilder)
```

```{r}
shopstatus <- installWorkshops("MultiAssayWorkshop",
    exclude = FALSE, ncpus = 12)
```

## Run `R CMD INSTALL *` on each repository

The output of the installation can be seen by using `getStatus`:

```{r}
getStatus()
```

## (optional) Post repository installation status

The text of the installation output can be posted as a comment in the
issue by running `postStatus`.

_Note_. This requires a `GITHUB_PAT` Personal Authentication Token (PAT)
environment variable (usually in `.Renviron`).

```{r}
postStatus()
```

## Inspect the `data.frame` with metadata for each workshop repository

This `data.frame` includes metadata including the issue number and
URL location

```{r}
shopstatus
```

## Capture the output of the installation and save to a file

Installation output is saved to a file in the `REPOS_PATH` folder; defaults
to "buildout".

```{r}
list.files(
    file.path(.options$get("REPOS_PATH"), "buildout")
)
```

## Clone the main workshop book repository

This can be done by running the `getBookRepo` function. It will take the
`BOOK_REPO` package value and clone from GitHub.

```{r}
getBookRepo()
```

## Add workshops to the DESCRIPTION file

Using the `desc` package, we then add all the successfully installed workshops
to the `DESCRIPTION` file of the book repository.

Notice that `shopstatus` has a logical column that indicates a successful
installation called 'install'.

```{r}
addWorkshops(shopstatus)
```

## Transfer vignettes to the main workshop book repository

We then transfer all successful workshop Rmd files to the main book repository
for building.

```{r}
rmds <- transferVignettes(shopstatus)
```

## Include the `Rmd` files in the configuration file for book building

We make sure to document the added R Markdown files in the `_bookdown.yml`
configuration file.

```{r}
addRmdFiles(rmds)
```

## Render the book

The configuration files that are pre-included in the workshop book repository
allow the book to build. In the future, we may consider adding this step
via the R interface. We use these files to render the book into the different
formats depending on the input to the `renderWorkshops` format.

```{r}
renderWorkshops(format = "all")
```
repo)

